manuscript

### 2025-04-29 doc/Manuscript.qmd
draft manuscript for circulation 04/26/2025

---
format:
  pdf:
    documentclass: article
    linestretch: 2
    number-sections: true
    include-in-header: draft-options.tex
bibliography: references.bib
execute: 
  echo: true
---

@@

John Imperato^1^

^1^Yale School of the Environment

@@

@@

Amphibians continue to experience global declines, with infectious diseases recognized as primary drivers of biodiversity loss [@luedtke2023a] (Luedtke et al. 2023). Chytridiomycosis, caused by the fungal pathogen *Batrachochytrium dendrobatidis* (Bd), has driven declines or extinctions of approximately 200 amphibian species globally [@skerratt2007] (Skerratt et al. 2007). Amphibian responses to Bd infection vary dramatically; some populations show minimal impact, others rapidly go extinct, and a rare few exhibit persistent recovery following severe declines [@lips2016] (Lips 2016). The mechanisms supporting persistence or recovery remain poorly understood but likely involve interactions among ecological, evolutionary, immunological, and demographic factors [@scheele2017] (Scheele et al. 2017).

The mountain yellow-legged (MYL) frog species complex (*Rana sierrae* and *R. muscosa*), endemic to California's Sierra Nevada and Transverse mountain ranges, provides a particularly well-documented case of amphibian decline and recovery from Bd infection. Starting in the 1970s, Bd spread through the Sierra Nevada, causing widespread and rapid extirpations of MYL frog populations [@vredenburg2010] [@vredenburg2019] (Vredenburg et al. 2010, 2019). By the early 2000s, over 90% of their historical range had been lost, due to combined impacts from Bd and habitat degradation caused by introduced trout [@knapp2000] (Knapp and Matthews 2000). Despite these losses, a few populations survived initial Bd outbreaks, eventually exhibiting increased abundance and apparent disease resistance [\@knapp2016](Knapp%20et%20al.%202016). These dynamics exemplify patterns of evolutionary rescue [@carlson2014] (Carlson et al. 2014), in which populations rapidly acquire adaptive genetic variation that restores demographic viability under persistent threats [@knapp2024] (Knapp et al. 2024).

While broad-scale studies have described these declines and recoveries, detailed demographic analyses at the individual population level remain scarce, primarily due to logistical challenges. Most fine-scale demographic research has focused on southern populations, particularly within Yosemite National Park [@fellers2013] [@joseph2018] [\@knapp2024](Fellers%20et%20al.%202013;%20Joseph%20and%20Knapp%202018;%20Knapp%20et%20al.%202024). In contrast, the population dynamics of frogs in the northern Sierra Nevada—where habitats are more heterogeneous and populations are smaller and lower in elevation—remain understudied [@brown2019] (MYLF ITT 2018; Brown et al. 2020).

To address this gap, we conducted a five-year capture-mark-recapture (CMR) study of a naturally recovering, Bd-persistent *R. sierrae* population in Tahoe National Forest. Our objectives were to estimate key demographic parameters—abundance, survival, recruitment, and detection probability—to provide insights into the ecological processes and factors that contribute to population viability under endemic Bd infection. We also examined the role of climate variability in shaping these demographic parameters, as climate change is increasingly recognized as a primary driver of amphibian declines worldwide, comparable in impact to disease and habitat loss [\@luedtke2023a](Luedtke%20et%20al.%202023). Understanding how climate influences abundance, survival, and recruitment is particularly important for high-elevation amphibian populations living with endemic Bd infection.

Understanding demographic patterns in recovering MYL frog populations has significant conservation implications. Recent research demonstrates these populations can serve as effective founders in reintroduction programs, establishing self-sustaining populations when translocated to suitable but currently unoccupied habitats [@knapp2024] (Knapp et al. 2024). Leveraging evolved disease resistance from naturally recovering populations offers a promising conservation strategy. Our study provides critical demographic data to support these conservation efforts, particularly for the less-studied northern Sierra Nevada populations.

@@

The Mossy Pond study area is a one square mile section of the Tahoe National Forest, north of Highway 80 in Nevada County, California. It is characterized by lakes, ponds, and streams set on granite benches, ranging from 6,400 feet to 7,100 feet in elevation. The study area includes the namesake Mossy Pond (6 hectare surface area, maximum depth 2.5 meters), its seasonally flowing outlet stream, and 12 ephemeral ponds. The absence of deep-water habitat is unusual for areas supporting MYL frog populations, given their reliance on water bodies deeper than 4 meters for overwinter survival [@bradford1983] [@knapp2003] (Bradford 1983; Knapp et al. 2003).

*R. sierrae* is a federally endangered, mid-sized ranid frog that occupies montane waterbodies in California's Sierra Nevada (U.S. Fish and Wildlife Service 2014). The US Geological Survey first detected *R. sierrae* in the Mossy Pond area in 1998, and the California Department of Fish and Wildlife (CDFW) has been monitoring the population since 2001. **More?**

@@

To assess the Mossy Pond *R. sierrae* population's Bd status, we collected epithelial swabs from 24 adult frogs over three years (2008, 2010, and 2021). Partner scientists quantified the amount of Bd DNA on each swab using real-time quantitative polymerase chain reaction (qPCR) analysis (Boyle et al. 2004). We report infection loads as the number of ITS1 (Internal Transcribed Spacer 1) gene copies detected [@longo2013] [@joseph2018] (Longo et al. 2013; Joseph and Knapp 2018). Quantifying Bd loads rather than simply testing for Bd presence allowed us to compare our study population with other known recovering populations. Partner scientists provided *R. sierrae* Bd load data from three known naturally recovering populations in Yosemite National Park [@knapp2024] (Knapp et al. 2024).

@@

We studied *R. sierrae* with capture-mark-recapture (CMR) methods, which are widely used for estimating wildlife population parameters such as abundance, survival, and recruitment (Williams et al. 2001). We conducted CMR sampling according to Pollock's robust design [@pollock1982], which involves sampling at two temporal scales: primary periods (between which the population is open to demographic change) and secondary periods, during which the population is assumed closed. Each summer from 2014 to 2018, we visited the study site 3 or 4 times (primary periods) for three consecutive days of surveys (secondary periods), for a total of 47 individual surveys.

Surveys occurred between 0800 and 1800, when *R. sierrae* are most active. We systematically searched all water bodies in the study area, capturing frogs by hand or dip net for immediate processing. We first scanned each captured frog with a BioMark 601 passive integrated transponder (PIT) tag reader (BioMark, Boise, Idaho, USA) to detect if the frog was marked (i.e., a recapture). We then used calipers to measure the snout to urostyle (SUL) length of each captured frog, and released frogs \<40 mm SUL (which are considered subadults) without further processing. Frogs larger than 40mm (adults) required further processing. New adult captures were PIT-tagged dorsally following McAlister et al. 2004 [@mcallister2004]. Next, we recorded the frog's sex and weighed it inside a tared plastic bag to the nearest 0.1 g using a Pesola spring scale. Finally, we collected a GPS point (estimated precision error \~3 m) for each capture using a handheld GPS unit. We only collected length and weight data during the first capture event within a three-day primary period. Afterward, if crews captured the same individual on a subsequent day during the same primary period, crews only recorded PIT tag, sex, and location data.

@@

We estimated survival, detection, recruitment, and adult population size for the Mossy Pond *R. sierrae* population using a site‐specific open‐population Jolly–Seber CMR model in a Bayesian framework with R statistical software \[[\@R](v4.4.2)\]. Our implementation builds on the mrmr R package [@joseph2019] (https://snarl1.github.io/mrmr/index.html), which tracks a superpopulation of M individuals through not‐recruited, alive, or dead states and models detection and demographic processes across primary and secondary periods [@knapp2024]. We assessed convergence with trace‐plots and Gelman–Rubin (R̂) diagnostics and compared model alternatives using leave‐one‐out cross‐validation.

@@

We defined winter severity as the percentage of historical average snow-water equivalent (SWE) remaining on April 1 each year (100% indicating average severity). We obtained SWE data from ten regional snow-survey stations (4.5–19.5 km from Mossy Pond) to capture the spatial variability in snowpack present in topographically complex mountainous environments [@revuelto2014] (Revuelto et al. 2014). Annual SWE values (2011–2018) were interpolated across the study area using inverse-distance weighting in ArcGIS Pro 3.4.3 (Esri Inc. 2025). From the resulting continuous SWE raster, we extracted the mean percent SWE within our one-square-mile study boundary. All raw SWE measurements were obtained from the California Data Exchange Center (CDEC; http://cdec.water.ca.gov).

@@

16 out of 24 samples tested positive for Bd, with infection intensities ranging from 16 to 43,318 ITS1 gene copies. These Bd loads are well below the \~600,000 gene copy threshold associated with severe chytridiomycosis and increased MYL frog mortality [@vredenburg2010] [@longo2013] (Vredenburg et al. 2010; Longo et al. 2013). The infection intensities observed in our study population are typical of those measured in other known naturally recovering *R. sierrae* populations [@knapp2024] (Figure 1; Knapp et al. 2024).

@@

The Mossy Pond frogs carried Bd at moderate prevalence (67%) but at infection intensities that were two to three orders of magnitude lower than levels associated with increased mortality in MYL frogs. Such low loads mirror those reported for other naturally recovering populations in Yosemite National Park and stand in stark contrast to the uniformly high loads (\>600,000 ITS1 copies) and near-certain mortality that typify collapsing populations [@vredenburg2010] (Vredenburg et al. 2010). Taken together with the population’s stable abundance over the study period, these data support the hypothesis that the Mossy Pond frogs have undergone an evolutionary rescue process in which genotypes conferring resistance have been favoured and now dominate the population. This expands the geographic scope of Bd-resistant *R. sierrae* beyond the southern Sierra Nevada, indicating that resistance is not confined to a single environmental setting or genetic lineage. **Too much?**

@@

Our Bayesian correlation analysis revealed a strong negative association between winter severity in the previous year and both adult recruitment and population size. Mild, low-snow winters may enhance juvenile overwinter survival and accelerate growth such that more animals reach the 40 mm adult threshold before the next breeding season. This mechanism has been invoked for long-lived amphibians in other montane systems [@mccaffery2010] (McCaffery & Maxell 2010) and is consistent with the lag-1 pattern we observe. The absence of any clear winter-severity signal in overwinter adult survival suggests that once frogs surpass a critical body size, they can tolerate a broad range of cold-season conditions.

@@

Monthly survival was consistently high during the four overwinter intervals (μ = 0.97) but dropped and became more variable in summer (μ = 0.80). Elevated summer mortality is probably multifactorial: frogs are active and conspicuous, water levels drop, and thermal and desiccation stress peak—all of which increase predation risk and physiological costs. Conversely, winter conditions at Mossy Pond may be unusually benign. Although the study area lacks the deep-water refuges (\>4m) traditionally thought essential, our capture records show adults concentrated in the rocky outlet stream and mud-bottomed main pond—habitats that can stay unfrozen and oxygenated thanks to possible groundwater inputs and insulating effects. **More from Isaac here?**

The pronounced male survival advantage across all intervals (Δφ = 0.02–0.07) contrasts with Fellers et al. (2013), who reported nearly identical annual survival of male and female R. sierrae at Summit Meadow, with females sometimes faring slightly better. Our pattern instead mirrors Rocky Mountain tailed frogs (Ascaphus montanus), where males have shown modestly higher survival [@honeycutt2019] (Honeycutt et al. 2019).

Several non-exclusive factors could explain why females at Mossy Pond survived slightly less well. Most obviously, the energetic burden of reproduction falls almost entirely on females: producing and depositing large gelatinous egg masses—often while bearing the added weight of one or more males in amplexus—depletes energy reserves and can delay post-breeding foraging. Post-oviposition condition declines have been linked to reduced annual survival in other anurans (e.g., common toads; [@madsen2010] Loman & Madsen 2010). Additionally, movement studies of *A. montanus* show that females travel farther than males to locate suitable oviposition sites, increasing exposure to predators and desiccating microhabitats [@honeycutt2019] (Honeycutt et al. 2019). Together, these reproductive and behavioural costs may magnify female summer mortality in the Mossy Pond *R. sierrae* population, driving the male-biased survival pattern we observed.

@@

Across the western United States, April-1 SWE has declined by about 18% on average since 1955, with particularly steep losses in northern California (US EPA 2024 indicator). Regional climate projections point to continued warming, more precipitation falling as rain rather than snow, and further reductions in spring snowpack over the coming decades [@gottlieb2024] (Gottlieb and Mankin 2024). In the short term, such trends could benefit Mossy Pond frogs by sustaining the sequence of “light” winters that boost recruitment. Over longer horizons, however, chronic snowpack loss and more frequent summer droughts are likely to lower water levels, shorten hydroperiods, and shrink the already shallow overwintering habitats available in the basin. The drop in visual-encounter counts recorded after the 2018–2021 drought hints that these longer-term drying forces may eventually override the positive recruitment signal.

@@

Our results place the Mossy Pond frogs squarely within the “resistant-donor” framework that Knapp et al. [@knapp2024] (2024) used to successfully facilitate landscape-scale *R. sierrae* recovery in Yosemite National Park. In that 15-year experiment, reintroductions sourced from naturally recovering, Bd-resistant *R. sierrae* populations established self-sustaining colonies at 75% of recipient sites despite the pathogen’s continued presence, and half of those new populations were projected to remain viable over 50 years. The Mossy Pond population’s large size, stable demographics, and demonstrated resistance make it a valuable additional conservation resource, especially for northern Sierra reintroductions where local adaptation may confer an advantage for long-term population viability.

@@

California Department of Water Resources. 2025. California Data Exchange Center (CDEC). California Department of Water Resources. <https://cdec.water.ca.gov>

Esri Inc. 2025. ArcGIS Pro (Version 3.4.3). Esri Inc. <https://www.esri.com/en-us/arcgis/products/arcgis-pro/overview>

MYLF ITT. 2018. USDA Forest Service, U.S. Fish and Wildlife Service, National Park Service, and California Department of Fish and Wildlife. *Interagency Conservation Strategy for Mountain Yellow-Legged Frogs in the Sierra Nevada*. U.S. Department of Agriculture, Forest Service, Pacific Southwest Region.

U.S. Environmental Protection Agency (EPA). (2024). Climate Change Indicators: Snowpack. Retrieved from <https://www.epa.gov/climate-indicators/climate-change-indicators-snowpack>

U.S. Fish and Wildlife Service. 2014. Endangered and threatened wildlife and plants; endangered species status for Sierra Nevada yellow-legged frog and northern distinct population segment of the mountain yellow-legged frog, and threatened species status for Yosemite toad. *Federal Register* 79(82):24256–24310.

Williams, B. K., Nichols, J. D., and Conroy, M. J. 2002. *Analysis and management of animal populations*. Academic Press, San Diego, California.

## Figures

>>>>

---
format:
  pdf:
    documentclass: article
    linestretch: 2
    number-sections: true
    geometry: margin=1in
    include-in-header: |
      \usepackage{lineno}
      \AtBeginDocument{\linenumbers}
bibliography: 
  - references.bib
  - manual.bib
execute:
  echo: false
---

@@

@@

Mountain yellow-legged frogs (*Rana sierrae* and *R. muscosa*) have disappeared from \>90% of their historic range due to *Batrachochytrium dendrobatidis* (Bd) infection, yet a handful of populations now persist and recover in the pathogen's presence. We studied one such recovering population at Mossy Pond, Tahoe National Forest, for five summers (2014–2018) using robust-design capture–mark–recapture methods and a Bayesian Jolly–Seber model. From 3,631 captures of 657 uniquely marked adults, estimated abundance varied from 117 to 245 individuals, while 629 new adults recruited over the study. Monthly survival was strongly seasonal—mean 0.97 in winter versus 0.80 in summer—with male survival exceeding female survival by 2–7% in each interval. Survey-specific detection probability averaged 0.45. Bd was present in 67% of skin-swab samples and infection loads remained two–three orders of magnitude below lethal thresholds. Recruitment and abundance declined sharply after severe winters, yet overwinter adult survival showed no snow effect. These findings confirm functional Bd resistance, reveal contrasting seasonal pressures that shape population dynamics, and identify this robust northern Sierra Nevada population as a prime donor for translocation-based recovery efforts.

@@

Amphibians continue to experience global declines, with infectious diseases recognized as primary drivers of biodiversity loss [@luedtke2023a]. Chytridiomycosis, caused by the fungal pathogen *Batrachochytrium dendrobatidis* (Bd), has driven declines or extinctions of approximately 200 amphibian species globally [@skerratt2007]. Amphibian responses to Bd infection vary dramatically; some populations show minimal impact, others rapidly go extinct, and a rare few exhibit persistent recovery following severe declines [@lips2016]. The mechanisms supporting persistence or recovery remain poorly understood but likely involve interactions among ecological, evolutionary, immunological, and demographic factors [@scheele2017].

The mountain yellow-legged (MYL) frog species complex (*Rana sierrae* and *R. muscosa*), endemic to California's Sierra Nevada and Transverse mountain ranges, provides a particularly well-documented case of amphibian decline and recovery from Bd infection. Starting in the 1970s, Bd spread through the Sierra Nevada, causing widespread and rapid extirpations of MYL frog populations [@vredenburg2010; @vredenburg2019]. By the early 2000s, over 90% of their historical range had been lost, due to combined impacts from Bd and habitat degradation caused by introduced trout [@knapp2000]. Despite these losses, a few populations survived initial Bd outbreaks, eventually exhibiting increased abundance and apparent disease resistance [@knapp2016]. These dynamics exemplify patterns of evolutionary rescue [@carlson2014], in which populations rapidly acquire adaptive genetic variation that restores demographic viability under persistent threats [@knapp2024].

While broad-scale studies have described these declines and recoveries, detailed demographic analyses at the individual population level remain scarce, primarily due to logistical challenges. Most fine-scale demographic research has focused on southern populations, particularly within Yosemite National Park [@fellers2013; @joseph2018; @knapp2024] . In contrast, the population dynamics of frogs in the northern Sierra Nevada—where habitats are more heterogeneous and populations are smaller and lower in elevation—remain understudied (MYLF ITT 2018) [@brown2019].

To address this gap, we conducted a five-year capture-mark-recapture (CMR) study of a naturally recovering, Bd-persistent *R. sierrae* population in Tahoe National Forest. Our objectives were to estimate key demographic parameters—abundance, survival, recruitment, and detection probability—to provide insights into the ecological processes and factors that contribute to population viability under endemic Bd infection. We also examined the role of climate variability in shaping these demographic parameters, as climate change is increasingly recognized as a primary driver of amphibian declines worldwide, comparable in impact to disease and habitat loss [@luedtke2023a]. Understanding how climate influences abundance, survival, and recruitment is particularly important for high-elevation amphibian populations living with endemic Bd infection.

Understanding demographic patterns in recovering MYL frog populations has significant conservation implications. Recent research demonstrates these populations can serve as effective founders in reintroduction programs, establishing self-sustaining populations when translocated to suitable but currently unoccupied habitats [@knapp2024]. Leveraging evolved disease resistance from naturally recovering populations offers a promising conservation strategy. Our study provides critical demographic data to support these conservation efforts, particularly for the less-studied northern Sierra Nevada populations.

@@

The Mossy Pond study area is a one square mile section of the Tahoe National Forest, north of Highway 80 in Nevada County, California. It is characterized by lakes, ponds, and streams set on granite benches, ranging from 6,400 feet to 7,100 feet in elevation. The study area includes the namesake Mossy Pond (6 hectare surface area, maximum depth 2.5 meters), its seasonally flowing outlet stream, and 12 ephemeral ponds. The absence of deep-water habitat is unusual for areas supporting MYL frog populations, given their reliance on water bodies deeper than 4 meters for overwinter survival [@bradford1983; @knapp2003].

*R. sierrae* is a federally endangered, mid-sized ranid frog that occupies montane waterbodies in California's Sierra Nevada [@u.s.fishandwildlifeservice2014]. The US Geological Survey first detected *R. sierrae* in the Mossy Pond area in 1998, and the California Department of Fish and Wildlife (CDFW) has been monitoring the population since 2001.

@@

To assess the Mossy Pond *R. sierrae* population's Bd status, we collected epithelial swabs from 24 adult frogs over three years (2008, 2010, and 2021). Partner scientists quantified the amount of Bd DNA on each swab using real-time quantitative polymerase chain reaction (qPCR) analysis [@boyle2007]. We report infection loads as the number of ITS1 (Internal Transcribed Spacer 1) gene copies detected [@longo2013; @joseph2018]. Quantifying Bd loads rather than simply testing for Bd presence allowed us to compare our study population with other known recovering populations. Partner scientists provided *R. sierrae* Bd load data from three known naturally recovering populations in Yosemite National Park [@knapp2024].

@@

We studied *R. sierrae* with capture-mark-recapture (CMR) methods, which are widely used for estimating wildlife population parameters such as abundance, survival, and recruitment [@williams2001]. We conducted CMR sampling according to Pollock's robust design [@pollock1982], which involves sampling at two temporal scales: primary periods (between which the population is open to demographic change) and secondary periods, during which the population is assumed closed. Each summer from 2014 to 2018, we visited the study site 3 or 4 times (primary periods) for three consecutive days of surveys (secondary periods), for a total of 47 individual surveys.

Surveys occurred between 0800 and 1800, when *R. sierrae* are most active. We systematically searched all water bodies in the study area, capturing frogs by hand or dip net for immediate processing. We first scanned each captured frog with a BioMark 601 passive integrated transponder (PIT) tag reader (BioMark, Boise, Idaho, USA) to detect if the frog was marked (i.e., a recapture). We then used calipers to measure the snout to urostyle (SUL) length of each captured frog, and released frogs \<40 mm SUL (which are considered subadults) without further processing. Frogs larger than 40mm (adults) required further processing. New adult captures were PIT-tagged dorsally following @mcallister2004. Next, we recorded the frog's sex and weighed it inside a tared plastic bag to the nearest 0.1 g using a Pesola spring scale. Finally, we collected a GPS point (estimated precision error \~3 m) for each capture using a handheld GPS unit. We only collected length and weight data during the first capture event within a three-day primary period. Afterward, if crews captured the same individual on a subsequent day during the same primary period, crews only recorded PIT tag, sex, and location data.

@@

We estimated survival, detection, recruitment, and adult population size for the Mossy Pond *R. sierrae* population using a site‐specific open‐population Jolly–Seber CMR model in a Bayesian framework with R statistical software [@R]. Our implementation builds on the mrmr R package [@joseph2019] (https://snarl1.github.io/mrmr/index.html), which tracks a superpopulation of M individuals through not‐recruited, alive, or dead states and models detection and demographic processes across primary and secondary periods [@knapp2024]. We assessed convergence with trace‐plots and Gelman–Rubin (R̂) diagnostics and compared model alternatives using leave‐one‐out cross‐validation.

@@

We defined winter severity as the percentage of historical average snow-water equivalent (SWE) remaining on April 1 each year (100% indicating average severity). We obtained SWE data from ten regional snow-survey stations (4.5–19.5 km from Mossy Pond) to capture the spatial variability in snowpack present in topographically complex mountainous environments [@revuelto2014]. Annual SWE values (2011–2018) were interpolated across the study area using inverse-distance weighting in ArcGIS Pro 3.4.3. From the resulting continuous SWE raster, we extracted the mean percent SWE within our one-square-mile study boundary. All raw SWE measurements were obtained from the California Data Exchange Center (CDEC; http://cdec.water.ca.gov).

@@

16 out of 24 samples tested positive for Bd, with infection intensities ranging from 16 to 43,318 ITS1 gene copies. These Bd loads are well below the \~600,000 gene copy threshold associated with severe chytridiomycosis and increased MYL frog mortality [@vredenburg2010; @longo2013]. The infection intensities observed in our study population are typical of those measured in other known naturally recovering *R. sierrae* populations [Figure 1; @knapp2024].

@@

The Mossy Pond frogs carried Bd at moderate prevalence (67%) but at infection intensities that were two to three orders of magnitude lower than levels associated with increased mortality in MYL frogs. Such low loads mirror those reported for other naturally recovering populations in Yosemite National Park and stand in stark contrast to the uniformly high loads (\>600,000 ITS1 copies) and near-certain mortality that typify collapsing populations [@vredenburg2010]. Coupled with the population’s stable abundance over the study period, these patterns are consistent with the idea that some degree of Bd resistance as arisen in this population. While such dynamics could reflect an evolutionary rescue process in which genotypes conferring lower susceptibility have increased in frequency, additional genetic and experimental evidence would be required to confirm that mechanism. This expands the geographic scope of Bd-resistant R. sierrae beyond the southern Sierra Nevada, indicating that resistance is not confined to a single environmental setting or genetic lineage.

@@

Our Bayesian correlation analysis revealed a strong negative association between winter severity in the previous year and both adult recruitment and population size. Mild, low-snow winters may enhance juvenile overwinter survival and accelerate growth such that more animals reach the 40 mm adult threshold before the next breeding season. This mechanism has been invoked for long-lived amphibians in other montane systems [@mccaffery2010] and is consistent with the lag-1 pattern we observe. The absence of any clear winter-severity signal in overwinter adult survival suggests that once frogs surpass a critical body size, they can tolerate a broad range of cold-season conditions.

@@

Monthly survival was consistently high during the four overwinter intervals (μ = 0.97) but dropped and became more variable in summer (μ = 0.80). Elevated summer mortality is probably multifactorial: frogs are active and conspicuous, water levels drop, and thermal and desiccation stress peak—all of which increase predation risk and physiological costs. Conversely, winter conditions at Mossy Pond may be unusually benign. Although the study area lacks the deep-water refuges (\>4m) traditionally thought essential, our capture records show adults concentrated in the rocky outlet stream and mud-bottomed main pond—habitats that can stay unfrozen and oxygenated thanks to possible groundwater inputs and insulating effects.

The pronounced male survival advantage across all intervals (Δφ = 0.02–0.07) contrasts with Fellers et al. (2013), who reported nearly identical annual survival of male and female R. sierrae at Summit Meadow, with females sometimes faring slightly better. Our pattern instead mirrors Rocky Mountain tailed frogs (*Ascaphus montanus*), where males have shown modestly higher survival [@honeycutt2019].

Several non-exclusive factors could explain why females at Mossy Pond survived slightly less well. Most obviously, the energetic burden of reproduction falls almost entirely on females: producing and depositing large gelatinous egg masses—often while bearing the added weight of one or more males in amplexus—depletes energy reserves and can delay post-breeding foraging. Post-oviposition condition declines have been linked to reduced annual survival in other anurans [e.g., common toads; @madsen2010]. Additionally, movement studies of *A. montanus* show that females travel farther than males to locate suitable oviposition sites, increasing exposure to predators and desiccating microhabitats [@honeycutt2019]. Together, these reproductive and behavioural costs may magnify female summer mortality in the Mossy Pond *R. sierrae* population, driving the male-biased survival pattern we observed.

@@

Across the western United States, April-1 SWE has declined by about 18% on average since 1955, with particularly steep losses in northern California [@usepa2016]. Regional climate projections point to continued warming, more precipitation falling as rain rather than snow, and further reductions in spring snowpack over the coming decades [@gottlieb2024]. In the short term, such trends could benefit Mossy Pond frogs by sustaining the sequence of “light” winters that boost recruitment. Over longer horizons, however, chronic snowpack loss and more frequent summer droughts are likely to lower water levels, shorten hydroperiods, and shrink the already shallow overwintering habitats available in the basin. The drop in visual-encounter counts recorded after the 2018–2021 drought hints that these longer-term drying forces may eventually override the positive recruitment signal.

@@

Our results place the Mossy Pond frogs squarely within the “resistant-donor” framework that @knapp2024 used to successfully facilitate landscape-scale *R. sierrae* recovery in Yosemite National Park. In that 15-year experiment, reintroductions sourced from naturally recovering, Bd-resistant *R. sierrae* populations established self-sustaining colonies at 75% of recipient sites despite the pathogen’s continued presence, and half of those new populations were projected to remain viable over 50 years. The Mossy Pond population’s large size, stable demographics, and demonstrated resistance make it a valuable additional conservation resource, especially for northern Sierra reintroductions where local adaptation may confer an advantage for long-term population viability.

@@
